\chapter{Definitions}

The basic terminology used in this thesis
was introduced by
\todo{Add Hammer.}
\citet{Schieber2005154},
\citet{Dubovsky2012}
and \citet{Husek2014, Husek2015}.
\todokucera{Dále používáte obecnou booleovskou terminologii, ke které by se hodila citace třeba té booleovský knížky od Y. Cramy a P.L. Hammera.
Crama, Hammer: Boolean Functions - Theory, Algorithms, and Applications}

\section{Vector operations}

Throughout the thesis we will use vectors
over small finite domains extensively.
We shall write a vector as a sequence of symbols,
for example $00101$ is a vector of length $5$.
The concatenation of two vectors $u$ and $v$
will be denoted simply as $uv$,
for example $00v11$ denotes a vector
which is formed as a concatenation
of the three vectors $00$, $v$ and $11$.

\begin{definition}
[Component extraction $\bit{v}{i}$
{\definitionsource{Schieber2005154}}] % [p.~156]
Let $v$ be a vector of length $n$ and $1 \leq i \leq n$.
Then $\bit{v}{i}$ is the $i$-th component of $v$.
\end{definition}

\begin{definition}
[Subsequence extraction $\bits{v}{a}{b}$
{\definitionsource{Schieber2005154}}] % [p.~157]
Let
$v$ be a vector of length $n$
and
$1 \leq a \leq b \leq n$.
Then $\bits{v}{a}{b}$ is the subvector of $v$
that starts at $a$-th position
and ends at $b$-th position
($\bits{v}{a}{b}
= \bit{v}{a} \bit{v}{a+1}
\ldots \bit{v}{b-1} \bit{v}{b}$).
\end{definition}

Notably,
$v = \bit{v}{1} \ldots \bit{v}{n} = \bits{v}{1}{n}$.

\begin{definition}
[Symbol repetition $\rep{\alpha}{n}$
{\definitionsource{Schieber2005154}}] % [p.~157]
Let $\alpha$ be a symbol (for example $0$ or $1$)
and $n \in \nat$.
Then $\rep{\alpha}{n}$ is the vector of length $n$
each component of which is equal to $\alpha$.
\end{definition}

For example
$\rep{0}{2} = 00$.

\section{Binary vectors}

\begin{definition}
[Binary vector
{\definitionsource{Schieber2005154}}] % [p.~156]
We will use the term \definiendum{binary vector}
to denote a vector over the Boolean domain.
That is, $x$ is a binary vector if and only if
$x \in \booldom^n$ for some $n \in \nat$.
We call such $x$ an \definiendum{$n$-bit binary vector}.
\end{definition}

Note that an $n$-bit binary vector $x$
represents the natural number
$\sum_{i = 1}^n \bit{x}{i} 2^{n-i}
= \sum_{i|\bit{x}{i}=1} 2^{n-i}$.
An $n$-bit vector
corresponds to a number between $0$ and $2^n - 1$.
We will not differentiate between a number
and its binary vector representation.
Specifically,
we will use the standard linear order of natural numbers
to compare binary vectors.
Note that the linear order on natural numbers
corresponds to lexicographic order
on binary vectors:
\begin{observation}
\label{observation:lexicographicorder}
If $a$ and $b$ are $n$-bit binary vectors,
then $a < b$
if and only if
there are binary vectors $c$,
$a'$ and $b'$
such that
$a = c 0 a'$ and $b = c 1 b'$.
We call such $c$ the \definiendum{longest common prefix}
of $a$ and $b$.
\end{observation}

\section{Boolean functions}

\begin{definition}[Boolean function]
\todo{Cite a source.}
$f: \booldom^n \rightarrow \booldom$ is
an \definiendum{$n$-ary Boolean function}.
\end{definition}

Since we will not be dealing with any non-Boolean functions,
in the remainder of the text,
we will use the terms
\quot{function} and \quot{Boolean function}
interchangeably.

\begin{definition}[True point]
\todo{Cite a source.}
Let $f$ be an $n$-ary Boolean function.
An $n$-bit binary vector $x$ is a \definiendum{true point}
of $f$ if and only if $f(x)  = 1$.
\end{definition}

Equivalently,
we define a \definiendum{false point} of $f$
as any point $x$
such that $f(x) = 0$.

We'll denote the set of all true points of $f$ as $TP(f)$
and the set of all false points as $FP(f)$.

\section{\texorpdfstring{\acrshort{dnf}}{DNF}
representations of Boolean functions}

% http://en.wikipedia.org/wiki/Boolean_function
Every $n$-ary Boolean function $f$ can be expressed
as a propositional formula $\mathcal{F}$ on $n$ variables
$x_1, \ldots, x_n$.
There's a natural bijection
between binary vectors of length $n$
and valuations of $n$ variables
($1$-bits in the binary vector
correspond exactly to $1$-valued variables
in the valuation).

% TODO: Add example Boolean function with logical representation etc.

% http://en.wikipedia.org/wiki/Disjunctive_normal_form
\begin{definition}[\Acrlong{dnf}]
\todo{Cite a source.}
% Dubovsky2012: Definitions 1.9-1.11, page 3
A \definiendum{literal} is an occurrence
of a variable ($x_i$) or its negation ($\nott{x_i}$)
in a propositional formula.
A \definiendum{term} is
an elementary conjunction of literals,
that is a conjunction in which every variable appears
at most once.
A \definiendum{\acrfull{dnf}} propositional formula
is a disjunction of terms.
We will often view terms as sets of literals
and \acrshort{dnf} formulas as sets of terms.
\end{definition}

% http://en.wikipedia.org/wiki/Valuation_%28logic%29
A \acrshort{dnf} formula $\mathcal{F}$ is
a \definiendum{\acrshort{dnf} representation}
of a Boolean function $f$
if and only if
$\fora{v \in \booldom^n}{f(v) = 1 \iff
\valuation{\mathcal{F}}{v} \equiv 1}$,
where $\valuation{\mathcal{F}}{v}$ is the formula
given by substituting
each occurrence of a variable $x_i$ in $\mathcal{F}$
with the value $\bit{v}{i}$
for every $i \in \curly{1, \ldots, n}$.

\begin{definition}[Minimum \acrshort{dnf} representation]
\todo{Cite a source.}
A \acrshort{dnf} representation of function $f$ is
\definiendum{minimum}
if and only if
there is no \acrshort{dnf} representation of $f$
with fewer terms.

We shall denote the size of
a minimum \acrshort{dnf} representation
of function $f$
by $\apply{dnf}{f}$.
\end{definition}

Note that a function may have more than one minimum
\acrshort{dnf} representation.

The focus of this thesis is
minimizing the number of terms
of \acrshort{dnf} representations
of certain Boolean functions.

\section{Ternary vectors}

\begin{definition}
[Ternary vector
{\definitionsource{Schieber2005154}}] % [p.~156]
A \definiendum{ternary vector}
is a vector over the set
$\{0, 1, \phi\}$.
\end{definition}

There's a natural bijection between ternary vectors
and \acrshort{dnf} terms
(elementary conjunctions of literals).
Each of the $n$ components of a ternary vector $T$
corresponds to an occurrence (or its absence)
of one of the $n$ variables of a term $C$:
\begin{center}
$$\begin{array}{cc}
\bit{T}{i} & C \cap \curly{x_i, \overline{x_i}} \\
\hline
\phi & \emptyset \\
1 & \curly{x_i} \\
0 & \curly{\overline{x_i}}
\end{array}$$
\end{center}

Note that since $C$ is elementary,
it can not contain both $x_i$ and $\overline{x_i}$
for any $i$.

Namely,
a \emph{binary} vector
(a ternary vector that does not contain any $\phi$)
corresponds to a full term
(that is one that contains every variable),
and the ternary vector $\rep{\phi}{n}$
corresponds to the empty term (tautology).
% TODO: Consider defining "more qualified", "most qualified", "least qualified".

It is easy to see that a set of ternary vectors,
each of length $n$,
corresponds to a \acrshort{dnf} \emph{formula}
on $n$ variables.

\begin{definition}
[Spanning
{\definitionsource{Schieber2005154}}] % [p.~156]
A ternary vector $T$ of length $n$ \definiendum{spans}
a binary vector $x$ of length $n$
if and only if
$\fora{i \in \curly{1, \ldots, n}}
{\bit{T}{i} = \phi \text{ or } \bit{T}{i} = \bit{x}{i}}$.

A set of ternary vectors $\mathcal{T}$
\definiendum{spans}
a binary vector $x$
if and only if
some vector in $\mathcal{T}$ spans $x$.
%$\ex{T \in \mathcal{T}}
%{T \text{ spans } x}$.
\end{definition}

The $i$-th symbol of a ternary vector
constrains the $i$-th symbol of the spanned binary vector
(or, equivalently, the valuation of the $i$-th variable).
If $\bit{T}{i}$ is a \quot{fixed bit}
(that is $\bit{T}{i} \in \curly{0, 1}$),
the spanned binary vector $x$
must have the same value
in its $i$-th position,
that is $\bit{x}{i} = \bit{T}{i}$.
If $\bit{T}{i}$ is the \quot{don't care symbol} $\phi$,
% "don't" is used in Schieber2005
the spanned binary vector
may have any value in its $i$-th position.

In the correspondence between ternary vectors and terms,
spanning corresponds to satisfying.
The ternary vector $T$ of length $n$ spans $x$
if and only if
$C(x) \equiv 1$,
where $C$ is the term on $n$ variables
that corresponds to $T$.
Similarly,
it is easy to see that
a set of ternary vectors $\mathcal{T}$ spans $x$
if and only if
$\mathcal{F}(x) \equiv 1$,
where $\mathcal{F}$ is the \acrshort{dnf} formula
on $n$ variables
that corresponds to $\mathcal{T}$.

\begin{definition}[Spanned set]
Let $T$ be a ternary vector of length $n$.
We denote the set of points spanned by $T$
as $span(T)$:

\begin{equation*}
span(T) = \curly{
x \in \booldom^n |
T\text{ spans }x
}
\end{equation*}
\end{definition}

Note that a ternary vector
with $m$ $\phi$-positions
spans $2^m$ binary vectors.

We'll also use a natural extension of spanning
to sets of ternary vectors --
if $\mathcal{T}$ is a set of ternary vectors, then

\begin{equation*}
span(\mathcal{T}) =
\bigcup_{T \in \mathcal{T}} span(T)
\end{equation*}

\begin{definition}[Exact spanning]
A ternary vector $T$ of length $n$
\definiendum{spans exactly}
a set of $n$-bit binary vectors $S$
if and only if
$span(T) = S$.

$T$ \definiendum{spans exactly}
an $n$-ary Boolean function $f$
if and only if
$span(T) = TP(f)$.
\end{definition}

Again,
we will generalize \definiendum{exact spanning}
to sets of ternary vectors
-- the set of ternary vectors $\mathcal{T}$
\definiendum{spans exactly} $S$
if and only if
$span(\mathcal{T}) = S$.

\begin{definition}
[Spanning set
{\definitionsource{Schieber2005154}}] % [p.~156]
Let $f$ be an $n$-ary Boolean function.
The set $\mathcal{T}$ of $n$-bit ternary vectors
is a \definiendum{spanning set} of $f$
if and only if
it $\mathcal{T}$ spans exactly $f$,
that is
$span(\mathcal{T}) = TP(f)$.
\end{definition}

In other words,
a spanning set of a function
spans all of its true points
and none of its false points.

Note that spanning sets of a function
correspond to \acrshort{dnf} representations
of the function.
The correspondence preserves size
(number of ternary vectors and terms, respectively),
so a minimum spanning set
corresponds to a minimum \acrshort{dnf} representation.

\begin{definition}
[Disjoint spanning set
{\definitionsource[Section 4]{Schieber2005154}}] % [p.~165]
A spanning set $\mathcal{T}$ is \definiendum{disjoint}
if and only if the spanned sets of its vectors
do not overlap,
that is
\[
\fora{T_i, T_j \in \mathcal{T}}{T_i = T_j \text{ or }
span(T_i) \cap span(T_j) = \emptyset}
\]
\end{definition}

In terms of \acrshort{dnf},
disjoint spanning sets
correspond to \acrshort{dnf} formulas
such that every valuation satisfies at most one term.

We have introduced two equivalent representations
of Boolean functions --
one based on ternary vectors
and the other being the \acrshort{dnf} representation.
\Cref{table:representations}
shows the corresponding terms side by side.
% Note the inconsistent use of "term" here and in the table
% caption.

\begin{table}[h]
\centering
\begin{tabular}{ll}
ternary vector & term \\
ternary vector set & \acrshort{dnf} formula \\
binary vector & variable valuation \\
ternary vector \emph{spans} a binary vector &
valuation \emph{satisfies} a term \\
ternary vector set \emph{spans} a binary vector &
valuation \emph{satisfies} a \acrshort{dnf} formula
\end{tabular}
\caption{
Corresponding terms used
in Boolean function representations
}
\label{table:representations}
\end{table}

\begin{definition}[Complement
{\definitionsource{Schieber2005154}}] % [p.~156]
\label{def:complement}
The complement of a binary symbol $\alpha$
($\alpha \in \booldom$),
denoted $\compl{\alpha}$,
is $1 - \alpha$.

We get the complement of a binary vector $x$
by flipping all of its bits,
that is $\compl{x}
= (\compl{x_1}, \ldots, \compl{x_n})
= \rep{1}{n} - x$.

The complement of the $\phi$ symbol is $\phi$
($\compl{\phi} = \phi$).

It follows that
we obtain the complement of a ternary vector
by flipping all of its fixed bits.

We generalize the notion of complement to sets of ternary vectors
-- if $\mathcal{T}$ is a set of ternary vectors,
then $\compl{\mathcal{T}} = \curly{\compl{T} | T \in \mathcal{T}}$.
\end{definition}

\section{Orthogonal sets and coverable functions}

Orthogonal set is a useful tool
for showing a lower bound on the size of all spanning sets of a function.
Orthogonal sets have been used both
by \citet{Schieber2005154} (implicitly)
and \citet{Dubovsky2012} (explicitly)
in the proofs of optimality of spanning algorithms.

\begin{definition}
[Orthogonality
{\definitionsource[p.~6]{Dubovsky2012}}] % Definition 1.15
\label{def:orthogonal}
Let $x$ and $y$ be true points of a Boolean function $f$
(that is $f(x) = f(y) = 1$).
The vectors $x$ and $y$ are \definiendum{orthogonal}
with respect to $f$
if and only if
every ternary vector that spans both $x$ and $y$
spans some false point of $f$.

A set of true points $S$ is \definiendum{orthogonal}
if and only if
all the vectors in $S$ are pairwise orthogonal.

We shall denote the size of
a maximum orthogonal set
of function $f$
with \definiendum{$\apply{ortho}{f}$}.
\end{definition}

\todomaybe[inline]{Get rid of collision, or prove that it is equivalent to orthogonality, preferably using established terminology.}
\todomaybe[inline]{Note that $x$ and $y$ collide on $z$ if and only if $z$ separates $x$ and $y$ in terminology of \citet[Definition 3.3]{Cepek2012}. Consider using the same terminology.}
\begin{definition}
[Collision]
\label{def:collision}
Let $x$, $y$ and $z$ be $n$-bit binary vectors.
We say that \definiendum{$x$ and $y$ collide
on the vector $z$}
if and only if
every ternary vector that spans both $x$ and $y$
necessarily also spans $z$.
\end{definition}

\begin{theorem}
[Collision and orthogonality]
Let $x$ and $y$ be true points of function $f$.
If the vectors $x$ and $y$
collide on some false point of $f$,
they are orthogonal.
\end{theorem}

\begin{proof}
The statement is a trivial application of
\cref{def:collision,def:orthogonal}.
\end{proof}

Note that the inverse implication holds as well
\citep[Lemma 3.7]{Cepek2012}
\todoforkucera{Does this Lemma 3.7 in \citet{Cepek2012} really prove this? (I believe it does.)}
but we will not use it in this thesis.
\todo{Do we use the theorem in the thesis? If not, throw it away, possibly along with definition of collision. If yes, label and cref it.}
\todomaybe{Add \quot{the proof is more complicated}. We would especially need some established terminology.}

\begin{theorem}
[$\apply{ortho}{f} \leq \apply{dnf}{f}$
{\theoremsource[Observation 1.1]{Dubovsky2012}} % p.~6
%{\theoremsource[Corollary 6.12]{Boros2010}}] % p.~89
{\theoremsource[Theorem 2.8, Corollary 3.2]{Cepek2012}}] % p.~368
\todomaybe{Cite \citet{Boros2010} instead of or along with \citet{Cepek2012}.}
\todomaybe{Cite Berge: Graphs and Hypergraphs, Theorem 5, pp. 420-42.}
\label{theorem:orthodnf}
For any Boolean function,
the size of its maximum orthogonal set
is at most as great
as the size of its minimum \acrshort{dnf} representation.
\end{theorem}

\begin{proof}
We will prove the statement by contradiction.
Let $f$ be a Boolean function such that
$\apply{ortho}{f} > \apply{dnf}{f}$.
Let $V$ be an orthogonal set of size $\apply{ortho}{f}$
and $\mathcal{T}$ be any spanning set
of size $\apply{dnf}{f}$
(such spanning set exists because of the correspondence
between \acrshort{dnf} representations and spanning sets).
Since $\size{\mathcal{T}} < \size{V}$
and $\mathcal{T}$ spans $TP(f) \supseteq S$,
there must be a ternary vector $T \in \mathcal{T}$
that spans
two distinct vectors $x, y \in V$.
However,
since $V$ is orthogonal,
$x$ is orthogonal to $y$,
so $T$ necessarily spans a false point,
which contradicts the premise that $\mathcal{T}$
is a feasible spanning set of $f$.
\end{proof}

\todoforkucera[inline]{Is the parallel between orthogonal vectors and essential sets presented in the following paragraph correct?}
Note that \citeauthor{Cepek2012}
show the inequality
for disjoint essential sets of implicants of $f$
in place of orthogonal vectors
\citep[Theorem 2.8]{Cepek2012}. %p.~368
An orthogonal set corresponds
to a set of pairwise disjoint
\emph{truepoint} essential sets
\citep[Section 3]{Cepek2012}, % p.~368
\todoforkucera{Does the correspondence really hold? Note that the essential sets only contain resolvents of prime implicates ($l(f)$), while we consider all ternary vectors (no matter what their relation to $f$ is).}
where truepoint essential set
is a special case of essential set.
\citeauthor{Cepek2012} continue to
show that the lower bound on $dnf(f)$ is maintained
if we restrict ourselves
to sets of truepoint essential sets
\citep[Corollary 3.2]{Cepek2012}. % p.~368
Also note that their results are presented in the dual
\todo{Is \quot{dual} a correct word to describe the relation?}
terminology of \acrfull{cnf},
as opposed to the \acrshort{dnf} terminology
used throughout this thesis.
\todomaybe{Add \quot{I have taken the liberty to translate.}.}

We conclude that the size of any orthogonal set
is a lower bound on the size of minimum
\acrshort{dnf} representation.
Specifically,
if we show a \acrshort{dnf} representation of a function
and an orthogonal set of the same size,
we conclude that the representation is minimum.
%(and the orthogonal set is maximum).
The orthogonal set certifies
the optimality of the \acrshort{dnf} representation.

\begin{definition}
[Coverability
{\definitionsource[Definition 2.9]{Cepek2012}}] % [p.~368]
\label{def:coverable}
Let $f$ be a Boolean function.
The function $f$ is \definiendum{coverable}
if and only if
$\apply{ortho}{f} = \apply{dnf}{f}$.
\end{definition}

Informally speaking,
a coverable function is one for which it may be
easy to prove optimality of a spanning set.
We will see
in \cref{chap:1interval,sec:2int2switch}
that all the $1$-interval
and $2$-switch $2$-interval functions
are coverable.
We will show examples of functions which are not coverable
in \cref{sec:3switch}.
\todoforkucera[inline]{Maybe remove or rephrase the preceding sentence since $k$-interval and $l$-switch functions are defined later.}

\section{\texorpdfstring{$k$}{k}-interval
Boolean functions}

The notation for $1$- and $2$-interval
functions was introduced by \citeauthor{Schieber2005154}
and \citeauthor{Dubovsky2012} respectively.
We shall generalize it to $k$-interval functions
for every $k \geq 0$.

\begin{definition}[$k$-interval Boolean function]
\label{def:kibf}
Let $n \geq 0$ and $k \geq 0$.
\todo{Update $k \geq 1$ to $k \geq 0$ in the rest of the text.}
Let $a_1, b_1, \ldots, a_k, b_k$ be $n$-bit binary vectors
such that $\rep{0}{n} \leq a_1 \leq b_1 < a_2
\leq \cdots < a_i \leq b_i < a_{i+1}
\leq \cdots < a_k \leq b_k \leq \rep{1}{n}$.
Then $\fnkab: \booldom{}^n \rightarrow \booldom{}$ is a function defined as follows:
$$
\fnkab(x) =
\begin{cases}
1 & \text{if } a_i \leq x \leq b_i \text{ for some } i \\
0 & \text{otherwise}
\end{cases}
$$

We call
$f^n_{\interval{a_1}{b_1}, \ldots, \interval{a_k}{b_k}}$
a \definiendum{$k$-interval Boolean function}
and the vectors $a_1, b_1, \ldots, a_k, b_k$ its
\definiendum{endpoints}.
\end{definition}

Note in particular that in case $n=0$,
the $1$-interval $0$-ary function
$\intervalfn{f}{0}{\epsilon}{\epsilon}$,
where $\epsilon$ denotes the binary vector of length $0$,
is properly defined and equivalent to tautology.
Similarly note that for $k=0$
a $0$-interval function $f^n_{\emptyset}$
is equivalent to contradiction.

Note that the inequalities ensure
that the intervals $\interval{a_i}{b_i}$ for all $i$s
are non-empty and do not intersect.
The inequalities, however, allow the intervals to be adjoint.
This means that a $k$-interval function
may be equivalent to a $k'$-interval function
for some $k' < k$.
We will often find it useful to use a less ambiguous notion of the number of intervals of a function:
\begin{definition}
[Proper $k$-interval Boolean function]
\label{def:properkibf}
Let
$f^n_{\interval{a_1}{b_1}, \ldots, \interval{a_k}{b_k}}$
be a $k$-interval Boolean function.
If the endpoints satisfy the inequalities
$b_1 < a_2 - 1$, \ldots, $b_i < a_{i+1} - 1$, \ldots,
$b_{k-1} < a_k - 1$
(in other words, adjacent intervals
are separated by at least one false point),
we call
$f^n_{\interval{a_1}{b_1}, \ldots, \interval{a_k}{b_k}}$
a \definiendum{proper $k$-interval Boolean function}.
\end{definition}

%Note that if $f$ is proper $k$-interval,
%it is also $k$-interval,
%and that if $f$ is $k$-interval
%and $g$ is \emph{proper} $l$-interval
%for $l > k$,
%then $f$ and $g$ differ.

Especially,
$\fnab$ is the function
whose true points form the interval $\interval{a}{b}$.

\Cref{tab:exampleintfns}
shows examples of interval functions.

\begin{table}[h]
\centering
\begin{tabular}{lll}
Interval function & Propositional formula & Description \\
\hline
%$\intervalfn{f}{2}{11}{11}$ &
%$x_1 \wedge x_2$ & Conjunction \\
%$\intervalfn{f}{2}{01}{11}$ &
%$x_1 \vee x_2$ & Disjunction \\
$\intervalfn{f}{n}{\rep{1}{n}}{\rep{1}{n}}$ &
$\bigwedge_i{x_i}$ & Conjunction \\
$\intervalfn{f}{n}{\rep{0}{n-1} 1}{\rep{1}{n}}$ &
$\bigvee_i{x_i}$ & Disjunction \\
$\intervalfn{f}{n}{\rep{0}{n}}{\rep{1}{n}}$ &
$1$ (an empty term) & Tautology \\
$\intervalfn{f}{n}{\rep{0}{n-1} 1}{\rep{1}{n-1} 0}$ &
$\nott{x_1 \ldots x_n} \wedge
\nott{\nott{x_1} \ldots \nott{x_n}}
\equiv \bigvee_{i \neq j}{x_i \nott{x_j}}$ &
Non-equivalence \\
$f^n_{\interval{\rep{0}{n}}{\rep{0}{n}},
\interval{\rep{1}{n}}{\rep{1}{n}}}$ &
$x_1 \ldots x_n \vee \nott{x_1} \ldots \nott{x_n}$ &
Equivalence
\end{tabular}
\caption{Examples of interval functions}
\label{tab:exampleintfns}
\end{table}

\begin{definition}
[$\intervalfns{k}$]
\label{def:intervalfns}
For any $k \geq 0$,
$\intervalfns{k}$ is the class
of all proper $k$-interval Boolean functions,
each of them represented by its endpoints.
\todoforkucera{Is this a correct definition? Isn't the representation requirement too vague?}
\end{definition}

\begin{definition}
[$\allintervalfns$]
\label{def:allintervalfns}
Let $\allintervalfns$ denote the class
of all $k$-interval functions for every $k \geq 0$,
each of them represented by its endpoints
($\allintervalfns = \bigcup_{k \geq 0}{\intervalfns{k}}$).
\end{definition}
\todo[inline]{Remove this definition if it isn't used in the text.}

\section{Minimization of \texorpdfstring{$k$}{k}-interval functions}

\begin{problem}
[Boolean minimization]
\label{problem:minimization}
Let $F$ be a class of Boolean functions.
Then the problem $\minimization{F}$ is defined as follows:
Given a Boolean function $f \in F$,
find a minimum spanning set of $f$.
\end{problem}

Having defined all the necessary terms,
let's present the central problem
we will discuss throughout the thesis:
\begin{problem}
[Minimization of proper $k$-interval functions]
\label{problem:allintervalfnsminimization}
The problem $\minimization{\allintervalfns}$ is defined as follows:
Given the endpoints
of a proper $k$-interval function $f$ for any $k \geq 0$,
find a minimum spanning set of $f$.
\end{problem}

Note that since the problem of Boolean minimization is not tractable \citep{Umans1998}
\todoforkucera{It's $\Sigma_2$-hard. Does that mean it's not tractable?}
and the class $\allintervalfns$ contains all Boolean functions,
we do not expect that we will be able to solve the problem $\minimization{\allintervalfns}$ in polynomial time.
However,
an algorithm polynomial with respect to $n$ still may exist and would be interesting to show.
\todoforkucera{Is there a $\Sigma_2$-hard problem that has a parametrized polynomial solution?}
\todo{Mention in Conclusion.}

We will also touch upon the alternative version
of the minimization problem
that involves disjoint representations:
\begin{problem}
[Disjoint minimization of $k$-interval function]
Given the endpoints
of a proper $k$-interval function $f$,
find a minimum disjoint spanning set of $f$.
\end{problem}

\section{\texorpdfstring{$l$}{l}-switch Boolean functions}

Given a $k$-interval function,
apparently the most important values are the endpoints of the intervals.
The important property of an interval endpoint is that
it is a place where the value of the function changes
(considering proper $k$-interval functions).
We can view these values as places of switch,
where $f$ switches its value from $0$ to $1$ or back.

The notion of switch was introduced by \citeauthor{Husek2014}\footnote{
\citeauthor{Husek2014} first used the Czech term
\quot{\foreignlanguage{czech}{zlom}}
in his thesis \citep[p.~13]{Husek2014}
and later translated the term as \quot{switch}
in personal communication \citep{Husek2015}.
}
and it turns out it is very useful when studying $2$-interval functions.

\begin{definition}
[$l$-switch function
{\definitionsource[pp.~13-14]{Husek2015}}] % Definitions 2.1, 2.3
Let $l \geq 0$.
A Boolean function $f$ is \definiendum{$l$-switch}
if and only if
there are exactly $l$ input vectors $x$ such that
$f(x) \neq f(x+1)$.

We call such $x$ a \definiendum{switch} of $f$.
\end{definition}

\begin{example}
[Switches and intervals]
\label{example:switchesintervals}
Depending on $f(0)$,
an $l$-switch Boolean function $f$
is proper $\ceil*{\frac{l}{2}}$-interval
in case $f(0) = 0$,
% f(0) = 0:
% 1-switch: 1-interval
% 2-switch: 1-interval
% 3-switch: 2-interval
% 4-switch: 2-interval
% l-switch: \ceil{l/2}-interval
or proper $\ceil*{\frac{l+1}{2}}$-interval
in case $f(0) = 1$.
% f(0) = 1:
% 1-switch: 1-interval
% 2-switch: 2-interval
% 3-switch: 2-interval
% 4-switch: 3-interval
% l-switch: \ceil{(l+1)/2}-interval

Equivalently,
a proper $k$-interval $n$-ary function is:
\begin{itemize}
\item
$(2k)$-switch in case $f(\rep{0}{n}) = f(\rep{1}{n}) = 0$
\item
$(2k-1)$-switch in case $f(\rep{0}{n}) \neq f(\rep{1}{n})$
\item
$(2k-2)$-switch in case $f(\rep{0}{n}) = f(\rep{1}{n}) = 1$
\end{itemize}
\end{example}

\todomaybe[inline]{Show that complementing bits preserves number of switches.}

\begin{definition}
[$\switchfns{l}$]
\label{def:switchfns}
For any $l \geq 0$,
$\switchfns{l}$ is the class
of all $l$-switch Boolean functions.
\todonote{Note that we don't specify any representation here. This is due to the fact that we only use $\switchfns{l}$ in intersection with $\intervalfns{k}$, which requires interval endpoint representation.}
\end{definition}
